---
title: "Week 3 homework assignment"
author: "Anubha Nagar"
date: "16th September 2021"
output: github_document
editor_options:
  chunk_output_type: inline
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DSI-EDA
## Professor Michael Shepherd

#This homework is designed to get you to review the in-class notes + r code as well as work on your own code.

## Part 1: Midwest Data

Recall our use of the `midwest` data from week3-day1. 

## 1. In your own words, what does the function in this line of "week3-day1.Rmd"  do?

```{r}
#g1 <- g1 + scale_y_continuous(breaks=seq(0, 1000000, 200000), labels = function(x){paste0(x/1000, 'K')})
```

Answer 1:

* Y axis is a continuous variable and hence we use scale_y_continuous

* The breaks refers to the y axis limits and the increment value. So in this example "breaks=seq(0, 1000000, 200000" means that the y axis starts from 0 and goes till 1000000 with increments of 200000. Each increment is plotted with a marker. 

* labels is used to make thousands in terms of 'K' for example 1000 is written as 1K and so on. So each y label is changed in terms of K. 

## 2. Starting with one variable: During week3-day1, we learned about how to make a scatterplot in ggplot using `midwest` data. This was a useful illustration for how to (1) make a guess at a bivariate relationship in the data and (2) explore it using a scatterplot. But ultimately the graphic wasn't that interesting. Sometimes we need to take a step back and simply plot one variable at a time. 

Explore the relationship of population totals by state. Include a clear title, and change the xlab and ylab to be easy to read words (labels), try using geom_col for this. Interestingly you *could* force the outcome using geom_histogram() but typically we want to use histograms for a singular variable.

```{r}
library(ggplot2)
library(tidyverse)
#Demographic information of midwest counties
# https://ggplot2.tidyverse.org/reference/midwest.html
data("midwest", package = "ggplot2")

midwest

```

```{r}
midwest %>%
  group_by(state)%>%
  summarize(total= sum(poptotal))%>%
  ggplot(aes(x = state, y = total))+ geom_col(aes(fill = state))+ ggtitle('Population of each State') + scale_y_continuous(breaks=seq(0, 12000000, 3000000), labels = function(x){paste0(x/1000000, 'M')})

```

## 3. Make a histogram  for the percent of people below poverty

```{r}
ggplot(midwest, aes(x = percbelowpoverty)) +
  geom_histogram(bins = 30, fill = "purple", alpha = 0.5)+labs(title = "Count below Poverty line")+scale_x_continuous(labels = function(x){paste0(x, '%')})

```

## Part 3

In class we worked on Nashville schools data. Print one `best` graphic from the Nashville schools data and write one paragraph about the graphic. If you are using the graphic your group made, try to improve it. As an added challenge for those who want one, create a completely different graphic. What did you learn? Why is this interesting? 

```{r}
# packages you need libraried for today
library(dplyr)
library(ggplot2)
metro_nash_schools <- read.csv("metro-nash-schools.csv")
head(metro_nash_schools)
```
# Is there a positive relationship between the number of economically disadvantaged students and english proficiency in public schools in Nashville? 


```{r}
g1 <- ggplot(metro_nash_schools, aes(x=Economically.Disadvantaged, y=Limited.English.Proficiency)) +
      geom_point(aes(col=School.Level)) + geom_smooth(method='lm', col='firebrick', size=0.5, se = F) + scale_colour_brewer(palette = 'Spectral') + labs(x = 'Economically Disadvantaged', y = 'English Proficiency', title='Economically Disadvantaged vs English Proficiency')
g1
```



